Can machine learning correct microwave humidity radiances for the influence of clouds?

### 2021-02-10 Article/Review_1/response_reviewer_1.tex

\documentclass[11pt,a4paper,draft]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{geometry}
\usepackage{xcolor}
%\usepackage{showframe} %This line can be used to clearly show the new margins

\newgeometry{vmargin={30mm}, hmargin={35mm,25mm}}

\begin{document}
\textbf{Response Reviewer 1}

\textbf{	Interactive comment on ``Can machine learning
	correct microwave humidity radiances for the
	influence of clouds?'' by Inderpreet Kaur et al.}

\begin{itemize}

\item 	precipitation and most dense clouds" I think the hydrometeor size is an important

\textcolor{blue}{Reply: We agree that the hydrometeor size is an important aspect, but when it comes to an overview of the contamination at 183\,GHz, precipitating clouds and clouds with large optical thickness have the strongest effect. Though a dense cloud could be composed of hydrometeors of different shapes and size. }

\item	L66: ``only using measurements (no background data involved)'' My understanding is
		that the method is in theory model-free, but for the demonstration in this study, sim-
		ulations (e.g. background simulated at MWHS2 frequency) are used. Am I correct?
		Maybe this should be stressed here. Why not try with real MWHS2 observation for the
		all sky dataset?

\textcolor{blue}{Reply: Yes, the demonstration of the entire concept is based on simulations. We do not try with the real MWHS-2 observations as for the training process we would still require the clear-sky simulations.}

\item
		L137 "Simulations for all three sensors are noise free, so to incorporate the measurement uncertainties, whenever needed, Gaussian noise is added according to the chan-
		nel NEDT (Table 1 ? Table 3)." Are the errors arising from the radiative transfer calculation accounted for?

\textcolor{red}{Reply: }

\item	
		L159: "for all selected quantile fractions " by quantile fraction, do you mean the n th
		amongst the 7 selected (from 0.2 to 99.8\%)? Also, 16, 50 and 85\% are not symmetric
		(rounding?)

\textcolor{blue}{Reply: Yes, the selected quantiles are the seven percentiles mentioned. We re-phrase some sentences here to make it more clear. 85\% is a typo, it should be 84\%. }

\item	
		L163: Pfreundschuh uses an indicator function I (=1 or 0) in the CRPS, the authors
		here use y, can they explain the difference?

\textcolor{blue}{Reply: In the study by Pfreundschuh et al., 2018, CRPS is defined as using an indicator function ($I$), however for QRNN the integral is evaluated only by approximation. To calculate the CRPS, we use a piece-wise linear fit to the approximate posterior CDF obtain from the predicted quantiles. This approximation is further used to calculate CRPS. }

\item	
		L173: " The input data is all-sky brightness temperature" the simulated one, even for
		MWHS2, right?

\textcolor{blue}{Reply: Yes, even for MWHS-2 we use simulated all-sky brightness temperatures.}

\item	L301: "0.15 K" should be -0.15

\textcolor{blue}{Reply: The correction is made.}

\item	
		L453: "Among these four channels, 150 GHz has the highest peaking function around 4
		C2km (Chen and Bennartz, 2020)" Could you please clarify what you mean here, 150GHz
		is neither the highest nor the lowest peaking channel nor it peaks at 4km. Channel 89,
		118+/-1.1, 118+/-2.5, and 150 GHz peak at 0.1, 9.6, 2.9, 1km, respectively (according
		to Chen and Bennartz, 2020), this can also be seen in Lawrence et al. (2018) Fig. 1
		through the Jacobians of channels 6 and 7.

\textcolor{blue}{Reply: The reviewer is correct. It was wrongly mentioned that 150 GHz is the highest peaking channel, at 4\,km. We have corrected the text and it now reads as:	
``The channels, 150\,GHz and 118.75$\pm$2.5 peak between surface and 4\,km. These channels can only provide coverage to the humidity channels in the lower and mid troposphere. However, the 183\,GHz channels are sensitive to hydrometeor content up to 10\,km. The channel 118.75$\pm$1.1 peaks around 10\,km, but such information is only partly relevant for the higher peaking channels of 183\,GHz.'' }

\item	L460: "but such information would be not be completely orthogonal" duplicate "be"

\textcolor{blue}{Reply: The duplicate word is removed. }

\item	
		Table 1: NEDT are constructor specifications, the real noise is lower, see
		Fig 5 Guo, Y., J. Y. He, S. Y. Gu, and N. M. Lu, 2019: Calibration and validation of
		Feng Yun-3-D microwave humidity sounder II. IEEE Geoscience and Remote Sensing
		Letters, doi:10.1109/LGRS.2019.2957403.	Tab 5 Carminati, F., Atkinson, N., Candy, B., Lu, Q.: Insights into the Microwave Instruments Onboard the Feng-Yun 3D Satellite: Data Quality and Assimilation in the Met Office NWP System. Adv. Atmos. Sci. (2020). https://doi.org/10.1007/s00376-020-0010-1

\textcolor{red}{Reply:  }

\end{itemize}

		\textbf{Questions:}
		\newline
\begin{itemize}

\item	
		Is this method applicable to IR e.g. to an ATOVS system?

\textcolor{blue}{Reply:  Yes, the method could potentially be applied to IR. The functioning and the performance of the cloud correction approach is based on the differences in cloud signatures over different frequencies in the same field of view. For IR, infact, this feature is regularly used to flag out cloud contamination. Most cloud flagging schemes for IR are based on brightness temperature threshold.
}

\item 		
		The authors explain that it could benefit the all-sky assimilation systems indirectly for
		the analysis increment. Instead (or in addition) could it be used to model the variable
		observation error (when and by how much to be inflated)? This would be, in my view,
		the most valuable.

\textcolor{blue}{Reply: }

\item	Could the uncertainty use for weak constraint 4dvar?

\textcolor{blue}{Reply:}

\item	
		What is the resource cost of this method (is this fast enough to be used in 1-h regional
		nwp with 30min window)?

\textcolor{blue}{Reply:}

\end{itemize}

\end{document}

### 2021-02-10 Article/Review_1/response_reviewer_2.tex

\documentclass[11pt,a4paper,draft]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{geometry}
\usepackage{xcolor}
%\usepackage{showframe} %This line can be used to clearly show the new margins

\newgeometry{vmargin={30mm}, hmargin={35mm,25mm}}

\begin{document}
\textbf{Response Reviewer 2}

\textbf{
	Interactive comment on ``Can machine learning
	correct microwave humidity radiances for the
	influence of clouds?'' by Inderpreet Kaur et al.}

\begin{itemize}

\item 			
			As I understand it, this study uses only simulated measurements (either with or without hydrometeors included), even for MWHS-2 for which actual measurements are available. This should be made clearer, especially given line 66, which states that no``background'' data is required for the method. I'd also like to see a bit more discussion 	about whether the results you present should be expected to hold for real data. An assumption that you are making is that your forward model can represent real clouds with enough fidelity that a model trained on simulated data will still work on real-world data. Can you point to any evidence to back up this assumption? Have you tried comparing the histograms of model-simulated Tbs with real-world MWHS-2 Tbs?

\textcolor{blue}{Reply: }

\item  	
			What is the computational cost of the QRNN method compared to simpler cloud-
			clearing filters? Could it feasibly be implemented into existing NWP models given
			current computational constraints?

\textcolor{blue}{Reply:}

\item  
			Lines 135-143: Why the large difference in number of cases simulated for ICI (220,000)
			vs. AWS (143,000), if they are both coming from the same population of CloudSat
			profiles? For the training dataset, you use about 75\% of total cases for MWHS-2,
			about 80\% of cases for ICI, and about 84\% of cases for AWS. Why this difference in
			proportions?

\textcolor{blue}{Reply: Yes, both ICI and AWS simulations come from same population of Cloudsat. For SMS, we had access to a smaller database, which was used for a earlier study. Due to difference in sizes of the available obseravtions, maximum possible observations were incorporated to allow  representation to each cloud event. }

\item  
			Line 340: ``With a test dataset of 70 000 samples, we cannot represent the far wings of the distribution accurately.'' Couldn't one apply this same reasoning to Figures 2,3, or
			8, which have density values below $10^{-4}$ included?

\textcolor{blue}{Reply: Yes, the same reasoning applies to Figures 2, 3 and 8 as well. We do not curtail the plot at $10^{-4}$ in these figures, so that it can be shown that most of the cases with density less than $10^{-4}$ correspond to low accuracy, due to their incomplete representation.}

\item 
			Figure 13: I wonder if there might be an easier, more concise way to evaluate whether
			the uncertainty intervals are properly calibrated. Namely, could you simply calculate
			how often the true value falls within the $\pm2\sigma$ uncertainty range, for each of the un-
			certainty bins ($0 - 3$K, $3 - 8$K, $8 +$ K) that you've included on the plot? If this percentage
			is significantly less than 95\%, it would suggests that your uncertainties are too small,
			while if it were closer to 100\% it would suggest your uncertainties are too large. Even
			if you don't get rid of Fig. 13, I think this would be useful information to include in the paper.

\textcolor{blue}{Reply:}

\item  
			Lines 498-499: ``. . . other underlying uncertainties not considered here.'' Whether here or elsewhere in the paper, I think you should talk a bit more about what these other uncertainties are (radiative transfer model errors should certainly be discussed), and what effects they might have on your results.

\textcolor{blue}{Reply:}

\end{itemize}

\textbf{Typos}
\begin{itemize}

\item Line 26: The phrase ``weather satellites are since some time equipped'' is confusing to
me. I suggest ``weather satellites have for some time been equipped. . .''

\textcolor{blue}{Reply: The sentence is re-written as per suggestion.}

\item Line 51: I believe you are missing the word ``on'' between ``predicated'' and ``Gaussian.''

\textcolor{blue}{Reply: The typo is corrected.}

\item Line 154: Should 85\% be changed to 84\%? That would be +2 sigma for a normal
distribution.

\textcolor{blue}{Reply: Yes, the predicted percentile should be 84\% instead of 85\%. The typo is corrected.}

\item Line 455:``. . . provide coverage to the humidity channels in, lower and mid troposphere''
This is confusing - get rid of the comma and add the word ``the'' instead?

\textcolor{blue}{Reply: The comma is removed and the word ``the'' is added.}

\item Line 494:	``QRNN predictions are weighted mean. . .'' I think this should say, ``QRNN
predictions are the weighted mean. . .''

\textcolor{blue}{Reply: The word ``the'' is added.}

\item Line 508: Missing ``the'' before ``same''

\textcolor{blue}{Reply: The word ``the'' is added.}

\end{itemize}

\end{document}

### 2021-02-10 Article/ici_submm_cloudcorr_v1.1.tex
started response to reviewer

\introduction
%
Satellite observations of humidity inside the troposphere are mainly performed
by downward-looking sensors. Among this class of observations, the frequency
range around 183\,GHz has a special position. Water vapour has a noticeable
transition at 22\,GHz, but it is relatively weak and only column values can be
derived \citep[e.g.,][]{schluessel1990atmospheric} for the observation geometry
of concern. The first transition in the microwave region that can be used to
derive altitude information, i.e.\ ``sounding'', is the one at
183\,GHz \citep{kakar1983retrieval,wang1983profiling}. On the other hand, at
infrared wavelengths a high number of water vapour transitions are found,
including some of high strength. As a consequence, infrared sounders can
provide humidity profiles with high precision and good vertical resolution, but
with strong limitations imposed by clouds. To be able to also sense humidity
inside and below clouds, weather satellites are since some time equipped with
channels around 183\,GHz. Today such channels are part of several sensors, such
as ATMS \citep[Advanced Technology Microwave Sounder,][]{weng2012introduction}.

@@

At 183\,GHz, the impact of hydrometeors typically causes a decrease in the observed radiance due to scattering from ice hydrometeors \citep[e.g.,][]{barlakas:three:20}. This implies that if any cloud contamination is missed by the filtering, a negative bias in the mean radiance, compared to the true clear-sky mean, may translate into a bias in humidity after the retrieval or assimilation. For NWP systems assimilating clear-sky observations, the effect of undetected clouds may be overcome by inflating the observational errors and diminishing the impact of observations. Furthermore, the mathematical assumptions of data assimilation (DA) are predicated Gaussian errors with no mean bias, and residual cloud impacts that cause a net bias are not easily handled by variational bias correction. One solution is to apply a very strict filtering, but this increases the rejection of clear-sky values, i.e.\ an important loss of useful data. Another limitation of existing filtering approaches is their ``one for all'' approach, i.e.\ observations in all 183\,GHz channels are either kept or rejected. This often rejects more observations than needed, as the channels differ in their altitude coverage. An observation could be cloudy in some channels and still the be clear-sky in others. To allow a channel specific filtering, data likely need to be combined in a more complex manner than simple differences, but it is unclear what type of regression would be best as the ideal solution would be scene-dependent. This points towards applying machine learning techniques \citep[e.g.,][]{favrichon2019detecting}. A maybe less obvious problem is the assignment of uncertainty to the filtered values. To our best knowledge, so far only estimates of mean and worst case errors exist in the literature. Some cases with relatively high cloud impact will likely be missed, while most cases are clear-sky from the start. As the remaining cloudy cases can cause significant biases, the likely solution is to apply a quite conservative (high) error estimate. However, this will unnecessarily downgrade the value of the truly clear-sky cases and the observations are used in a non-optimal manner.

@@

For this application, the predicted percentiles are chosen to be
$0.2\,\%, 3\,\%, 16\,\%, 50\,\%, 85\,\%, 97\,\%$ and $99.8\,\%$. For a Gaussian
distribution with mean $\mu$ and standard deviation $\sigma$, these quantiles
approximately correspond to $\mu -3\sigma, \mu-2\sigma, \mu-\sigma
, \mu, \mu + \sigma, \mu + 2\sigma, \mu + 3\sigma$ and thus allows
estimation of the $\pm\, 1, \pm\, 2$ and $\pm\, 3\sigma$ confidence intervals.

QRNN's are trained to minimize the mean of the sum of the quantile loss functions,
%
\begin{align}
\mathcal{L}_\tau(y_\tau, y) &=
\begin{cases}
\tau|y - y_\tau|, & y_\tau < y \\ (1 - \tau)|y_\tau - y|, & \text{otherwise}
\\
\end{cases}
\end{align}
%
for all selected quantile fractions $\tau$, where $y_\tau$ is the predicted
quantile and $y$ the reference value from the training or test data. The
quantile loss is also used in this study as a performance criterion for the tuning of
the hyper-parameters of the QRNN (see Appendix~\ref{appendix:hyperparamter}). In
addition to the quantile loss, also the Continuously Ranked Probability Score
(CRPS) is considered. Given a predicted CDF
$F$ and the reference value $y$, the CRPS is defined as
%
\begin{align}
\text{CRPS}(F, y) = \int_{-\infty}^{\infty} \left (F(y') - y\right )^2\: dy'.
\end{align}
%
To compute the CRPS for a prediction from a QRNN, the predicted quantiles are
used to derive a piece-wise linear approximation of the CDF of the predicted
distribution. Note that CRPS is only used to evaluate hyper-parameter tuning.

@@

For channel 11, the bias after correction is $-$0.12\,K in comparison to 0.15\,K in the all-sky simulations. The decrease in bias is not significantly high, but the strong negative tail  diminishes after correction, indicating removal of cases with large deviations. Nonetheless, the non-zero negative skewness also indicates presence of cases with residual cloud impact in the corrected dataset. Filtering the cases with high cloud impact has only a marginal positive effect. A similar performance is evident for channel 12. Correction reduces the bias to $-$0.12\,K from $-$0.29\,K and standard deviation to 1.08\,K from 2.02\.K. The MAE is approximately 16\,\% lower after correction, but still the negative skewness is not removed completely. For channel 13 and 15 also, a similar performance is seen. But in the latter, error distributions are more symmetric and have the largest spread as compared to other four channels. The effect of poor predictions is highest in channel 15 owing to its maximum sensitivity to hydrometeor impact.

@@

The partial performance of QRNN is due to incomplete complementary information to 183\,GHz channels. The weighting functions of window channels 89 and 150\,GHz, and the two 118\,GHz channels peak in the lower troposphere. Among these four channels, 150\,GHz has the highest peaking function around 4\,km \citep{chen2020mwhs}. These channels can only provide coverage to the humidity channels in, lower and mid troposphere; nonetheless, the 183\,GHz channels are sensitive to hydrometeor content up to 10\,km. Due to missing complementary information from other channels in the upper troposphere, QRNN fails at predicting these cases accurately. Such cases are mostly associated with thin cirrus clouds, which have very small influence at 89 and 150\,GHz. Without additional information from other channels we cannot expect QRNN to perform better. Among the other available channels, the overlapping weighting functions of 183\,GHz can provide auxiliary information to train QRNN, but such information would be not be completely orthogonal or statistically uncorrelated. Results show that these channels indeed help in improving the training, yet non-orthogonality introduces highly correlated observational errors. The correlations for cloudy observations are not surprising as the cloud amount for different channels depends upon each other in a systematic way. However for clear-sky observations, the correlations should preferably be close to zero. This is observed to be true with QRNN-single, but QRNN-all fails at preserving the noise stochasticity. In the absence of hydrometeor impact, all 183\,GHz channels provide same information to the learning model, introducing redundancy. Redundant patterns in machine learning models often have undesirable effect on the predictive performance. In our application, redundant information does not affect the prediction accuracy, but introduces highly correlated observational errors between channels. Correlated errors are also undesirable for DA systems, but in future, if the operational centres progress with approaches dealing with correlated observation errors, for example as currently done for ATMS at ECMWF \citep{Weston2018eumATMS}, concurrent use of 89, 150 and 183\,GHz channels would give the best cloud correction performance.

@@

With ICI and SMS, for some channels the variability of errors smaller than measurement noise is achieved. This is a consequence of predictions for cases which lack cloud impact. QRNN predictions are weighted mean of measurements between channels. In the absence of clouds, also the sub-mm channels provide humidity information that is incorporated in the 183 GHz NFCS estimate and some compensation of noise can be achieved. This effect is observed to be stronger in ICI than SMS, as the former has a higher number of channels giving redundant information. Note that with actual satellite measurements, the spread of error distributions smaller than sensor noise could be difficult to achieve due to other underlying uncertainties not considered here.

@@

Another advantage of QRNN is the estimation of case-specific uncertainties. That is, the predictions over chosen quantiles quantify the underlying uncertainty of the particular case, and not just represent some ensemble mean error. This has the consequence that a DA system can assign a proper weighting of each individual QRNN prediction. The analysis of QRNN predicted errors and calibration plots confirmed that QRNN is successful in providing well calibrated probabilistic predictions also in practice, except for few cases associated with high error. Poorly calibrated predictions are a consequence of outliers which are not well represented in the training dataset. The a priori distribution of the training dataset is dominated by clear-sky cases, however, the cloudy cases which occur infrequently or lack independent complementary information cannot be represented by same a priori distribution. The distribution of rare cases can be improved by increasing the training dataset size, but lack of complementary information can only be balanced by including additional training inputs, e.g. brightness temperatures for other channels.

>>>>

\introduction
%
Satellite observations of humidity inside the troposphere are mainly performed
by downward-looking sensors. Among this class of observations, the frequency
range around 183\,GHz has a special position. Water vapour has a noticeable
transition at 22\,GHz, but it is relatively weak and only column values can be
derived \citep[e.g.,][]{schluessel1990atmospheric} for the observation geometry
of concern. The first transition in the microwave region that can be used to
derive altitude information, i.e.\ ``sounding'', is the one at
183\,GHz \citep{kakar1983retrieval,wang1983profiling}. On the other hand, at
infrared wavelengths a high number of water vapour transitions are found,
including some of high strength. As a consequence, infrared sounders can
provide humidity profiles with high precision and good vertical resolution, but
with strong limitations imposed by clouds. To be able to also sense humidity
inside and below clouds, weather satellites have for sometime been equipped with
channels around 183\,GHz. Today such channels are part of several sensors, such
as ATMS \citep[Advanced Technology Microwave Sounder,][]{weng2012introduction}.

@@

At 183\,GHz, the impact of hydrometeors typically causes a decrease in the observed radiance due to scattering from ice hydrometeors \citep[e.g.,][]{barlakas:three:20}. This implies that if any cloud contamination is missed by the filtering, a negative bias in the mean radiance, compared to the true clear-sky mean, may translate into a bias in humidity after the retrieval or assimilation. For NWP systems assimilating clear-sky observations, the effect of undetected clouds may be overcome by inflating the observational errors and diminishing the impact of observations. Furthermore, the mathematical assumptions of data assimilation (DA) are predicated on Gaussian errors with no mean bias, and residual cloud impacts that cause a net bias are not easily handled by variational bias correction. One solution is to apply a very strict filtering, but this increases the rejection of clear-sky values, i.e.\ an important loss of useful data. Another limitation of existing filtering approaches is their ``one for all'' approach, i.e.\ observations in all 183\,GHz channels are either kept or rejected. This often rejects more observations than needed, as the channels differ in their altitude coverage. An observation could be cloudy in some channels and still the be clear-sky in others. To allow a channel specific filtering, data likely need to be combined in a more complex manner than simple differences, but it is unclear what type of regression would be best as the ideal solution would be scene-dependent. This points towards applying machine learning techniques \citep[e.g.,][]{favrichon2019detecting}. A maybe less obvious problem is the assignment of uncertainty to the filtered values. To our best knowledge, so far only estimates of mean and worst case errors exist in the literature. Some cases with relatively high cloud impact will likely be missed, while most cases are clear-sky from the start. As the remaining cloudy cases can cause significant biases, the likely solution is to apply a quite conservative (high) error estimate. However, this will unnecessarily downgrade the value of the truly clear-sky cases and the observations are used in a non-optimal manner.

@@

For this application, the quantile fractions $\tau$ are chosen as 
$0.2\,\%, 3\,\%, 16\,\%, 50\,\%, 84\,\%, 97\,\%$ and $99.8\,\%$ percentiles. For a Gaussian
distribution with mean $\mu$ and standard deviation $\sigma$, these quantiles
approximately correspond to $\mu -3\sigma, \mu-2\sigma, \mu-\sigma
, \mu, \mu + \sigma, \mu + 2\sigma, \mu + 3\sigma$ and thus allows
estimation of the $\pm\, 1, \pm\, 2$ and $\pm\, 3\sigma$ confidence intervals.

QRNN's are trained to minimize the mean of the sum of the quantile loss functions,
%
\begin{align}
\mathcal{L}_\tau(y_\tau, y) &=
\begin{cases}
\tau|y - y_\tau|, & y_\tau < y \\ (1 - \tau)|y_\tau - y|, & \text{otherwise}
\\
\end{cases}
\end{align}
%
for all selected quantile fractions, where $y_\tau$ is the predicted
quantile and $y$ the reference value from the training or test data. The
quantile loss is also used in this study as a performance criterion for the tuning of
the hyper-parameters of the QRNN (see Appendix~\ref{appendix:hyperparamter}). In
addition to the quantile loss, also the Continuously Ranked Probability Score
(CRPS) is considered. Given a predicted CDF
$F$ and the reference value $y$, the CRPS is defined as
%
\begin{align}
\text{CRPS}(F, y) = \int_{-\infty}^{\infty} \left (F(y') - y\right )^2\: dy'.
\end{align}
%
To compute the CRPS for a prediction from a QRNN, the predicted quantiles are
used to derive a piece-wise linear approximation of the CDF of the predicted
distribution. Note that CRPS is only used to evaluate hyper-parameter tuning.

@@

For channel 11, the bias after correction is $-$0.12\,K in comparison to $-$0.15\,K in the all-sky simulations. The decrease in bias is not significantly high, but the strong negative tail  diminishes after correction, indicating removal of cases with large deviations. Nonetheless, the non-zero negative skewness also indicates presence of cases with residual cloud impact in the corrected dataset. Filtering the cases with high cloud impact has only a marginal positive effect. A similar performance is evident for channel 12. Correction reduces the bias to $-$0.12\,K from $-$0.29\,K and standard deviation to 1.08\,K from 2.02\.K. The MAE is approximately 16\,\% lower after correction, but still the negative skewness is not removed completely. For channel 13 and 15 also, a similar performance is seen. But in the latter, error distributions are more symmetric and have the largest spread as compared to other four channels. The effect of poor predictions is highest in channel 15 owing to its maximum sensitivity to hydrometeor impact.

@@

The partial performance of QRNN is due to incomplete complementary information to 183\,GHz channels. The weighting functions of window channels 89 and 150\,GHz, and 118$\pm$2.5\,GHz channels peak in the lower troposphere. The channels, 150\,GHz and 118.75$\pm$2.5 peak between surface and 4\,km. These channels can only provide coverage to the humidity channels in the lower and mid troposphere. However, the 183\,GHz channels are sensitive to hydrometeor content up to 10\,km. The channel 118.75$\pm$1.1 peaks around 10\,km, but such information is only partly relevant for the higher peaking channels of 183\,GHz. Due to missing complementary information from other channels in the upper troposphere, QRNN fails at predicting these cases accurately. Such cases are mostly associated with thin cirrus clouds, which have very small influence at 89 and 150\,GHz. Without additional information from other channels we cannot expect QRNN to perform better. Among the other available channels, the overlapping weighting functions of 183\,GHz can provide auxiliary information to train QRNN, but such information would not be completely orthogonal or statistically uncorrelated. Results show that these channels indeed help in improving the training, yet non-orthogonality introduces highly correlated observational errors. The correlations for cloudy observations are not surprising as the cloud amount for different channels depends upon each other in a systematic way. However for clear-sky observations, the correlations should preferably be close to zero. This is observed to be true with QRNN-single, but QRNN-all fails at preserving the noise stochasticity. In the absence of hydrometeor impact, all 183\,GHz channels provide the same information to the learning model, introducing redundancy. Redundant patterns in machine learning models often have undesirable effect on the predictive performance. In our application, redundant information does not affect the prediction accuracy, but introduces highly correlated observational errors between channels. Correlated errors are also undesirable for DA systems, but in future, if the operational centres progress with approaches dealing with correlated observation errors, for example as currently done for ATMS at ECMWF \citep{Weston2018eumATMS}, concurrent use of 89, 150 and 183\,GHz channels would give the best cloud correction performance.

@@

With ICI and SMS, for some channels the variability of errors smaller than measurement noise is achieved. This is a consequence of predictions for cases which lack cloud impact. QRNN predictions are the weighted mean of measurements between channels. In the absence of clouds, also the sub-mm channels provide humidity information that is incorporated in the 183 GHz NFCS estimate and some compensation of noise can be achieved. This effect is observed to be stronger in ICI than SMS, as the former has a higher number of channels giving redundant information. Note that with actual satellite measurements, the spread of error distributions smaller than sensor noise could be difficult to achieve due to other underlying uncertainties not considered here.

@@

Another advantage of QRNN is the estimation of case-specific uncertainties. That is, the predictions over chosen quantiles quantify the underlying uncertainty of the particular case, and not just represent some ensemble mean error. This has the consequence that a DA system can assign a proper weighting of each individual QRNN prediction. The analysis of QRNN predicted errors and calibration plots confirmed that QRNN is successful in providing well calibrated probabilistic predictions also in practice, except for few cases associated with high error. Poorly calibrated predictions are a consequence of outliers which are not well represented in the training dataset. The a priori distribution of the training dataset is dominated by clear-sky cases, however, the cloudy cases which occur infrequently or lack independent complementary information cannot be represented by the same a priori distribution. The distribution of rare cases can be improved by increasing the training dataset size, but lack of complementary information can only be balanced by including additional training inputs, e.g. brightness temperatures for other channels.
